### 2024-11-21 paper/paper.qmd
update graphs and tables for data / result section

---
title: "How Drug Usage During Labor and Delivery Affects Infant Health: A Predictive Analysis of APGAR5 Scores"
subtitle: "!!! MAIN RESULTS !!!"
author: 
  - Yunkyung Ko
thanks: "Code and data are available at: [https://github.com/koyunkyung/infant_health](https://github.com/koyunkyung/infant_health)."
date: today
date-format: long
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
format: pdf
number-sections: true
bibliography: references.bib
toc: true
toc-title: "Table of Contents"
toc-depth: 2
toc-location: left
---

```{r}
#| include: false
#| warning: false
#| message: false

library(tidyverse)
library(data.table)
library(arrow)
library(scales)
library(dplyr)
library(knitr)

raw_data <- fread("../data/01-raw_data/infant_data.csv" , showProgress = FALSE)
analysis_data <- read_csv("../data/02-analysis_data/infant_health.csv")
balanced_data <- read_csv("../data/02-analysis_data/infant_balanced.csv")

```

# Introduction

Overview paragraph

Estimand paragraph

Results paragraph

Why it matters paragraph

Telegraphing paragraph: The remainder of this paper is structured as follows. @sec-data....

# Data {#sec-data}

## Overview

We use the statistical programming language R [@citeR].... Our data [@shelter].... Following @tellingstories, we consider...

Overview text

## Measurement

Some paragraphs about how we go from a phenomena in the world to an entry in the dataset.

## Outcome variables

### Apgar Score: a measure the infant's chance of surviving the first year of life

The Apgar score is a measure of the need for resuscitation [@natality2023] to the infant, which is the act of bringing someone back to life or waking them [@cambridgeResuscitation]. It is a test given to newborns soon after birth (5 minutes) to check 'Appearance(skin color)', 'Pulse(heart rate)', 'Grimace response(reflexes)', 'Activity(muscle tone)', 'Respiration(breathing rate and effort)' [@kidsHealthApgar]. Each is rated on a scale of 0 to 2, with 2 being the best score [@kidsHealthApgar]. Apgar scores range from 0 to 10, with a score of 7 or higher indicating that the neonate is in good to excellent physical condition [@natality2023].

@fig-apgar1 shows that the majority of infants in the raw dataset achieve a high APGAR5 score, clustering around 9 and 10. Very few observations exist for lower scores, reflecting rare instances of significant distress at birth. Even after filtering the dataset by selecting the relevant variables for analysis and removing the NA values, @fig-apgar2 shows that the observations are overly clustered around high APGAR5 scores.

```{r}
#| label: fig-apgar1
#| fig-cap: "The distribution of APGAR5 scores across the entire observation in the original dataset **Note:** Unknown or unreported observations were excluded so that the distribution could be clearly visualized"
#| echo: false
#| warning: false
#| message: false

# Remove rows where 'apgar5' equals 99
filtered_data <- raw_data %>%
  filter(apgar5 != 99)

# Histogram for 'apgar5'
filtered_data |>
  ggplot(aes(x = apgar5)) +
  geom_histogram(aes(y = after_stat(count)), 
    binwidth = 1, 
    color = "grey", 
    fill = "skyblue", 
    alpha = 0.7) +

  # Add labels
  labs(x = "APGAR5 Score",
       y = "Count of Observations") +

  # Format y-axis numbers with commas
  scale_y_continuous(labels = comma) +

  # Apply a clean theme
  theme_minimal()

```

```{r}
#| label: fig-apgar2
#| fig-cap: The distribution of APGAR5 scores across the filtered observation in analysis dataset, which selected the relevant variables and removed the NA values
#| echo: false
#| warning: false
#| message: false

# Histogram for 'apgar5'
analysis_data |>
  ggplot(aes(x = apgar5)) +
  geom_histogram(aes(y = after_stat(count)), 
    binwidth = 1, 
    color = "grey", 
    fill = "skyblue", 
    alpha = 0.7) + 
  # Add labels
  labs(x = "APGAR5 Score",
       y = "Count of Observations") +
  # Format y-axis numbers with commas
  scale_y_continuous(labels = comma) +
  # Apply a clean theme
  theme_minimal()

```

To ensure a balanced analysis and accurately assess the effects of drug treatment during delivery and labor, the data was refined to achieve a more even distribution of observations across APGAR5 scores. Based on the lowest observation count of 2,065 in the original distribution, the number of observations for each APGAR5 score was set to 2,000 as shown in @fig-apgar3. Random sampling was used to ensure a fair distribution across all score levels.

```{r}
#| label: fig-apgar3
#| fig-cap: The distribution of APGAR5 scores across the filtered observation in analysis dataset
#| echo: false
#| warning: false
#| message: false

# Generate a table for APGAR5 score distribution
apgar_table <- balanced_data %>%
  group_by(apgar5) %>%  # Group by APGAR5 score
  summarise(
    Count = n()  # Count of observations
  ) %>%
  arrange(apgar5)  # Ensure rows are ordered by APGAR5 score

# Transpose the table for horizontal display
apgar_table_transposed <- t(as.matrix(apgar_table$Count))  # Transpose only the counts
colnames(apgar_table_transposed) <- apgar_table$apgar5  # Set APGAR5 scores as column names
rownames(apgar_table_transposed) <- "Count"  # Set the row name

# Convert to a data frame for display
apgar_table_transposed_df <- as.data.frame(apgar_table_transposed)

# Display the transposed table using kable
knitr::kable(
  apgar_table_transposed_df)

```

## Predictor variables

All of the predictor variables used in the analysis are classified into the same category, which is the 'characteristics of labor and delivery'. This item, which contains 6 separate checkboxes that the respondent can choose from, allows for the reporting of more than one characteristic and includes a choice of "None of the above"[@natality2023].

### Number of Treatments Used During Delivery and Labor

@fig-numtreat shows that the majority of observations involve 0 to 2 treatments during labor and delivery, with a steep decline in counts for 3 or more treatments. Most births reported in this dataset occur with minimal medical intervention, and higher number of treatments are relatively rare.

```{r}
#| label: fig-numtreat
#| fig-cap: The distribution of observations across different number of treatments used during delivery and labor
#| echo: false
#| warning: false
#| message: false

anal_data <- balanced_data %>%
  mutate(
    drug_count = rowSums(select(., c(indc, augmt, ster, antb, chor, anes)) == 1, na.rm = TRUE)
  )

# Calculate observation counts for each drug_count
drug_count_summary <- anal_data %>%
  group_by(drug_count) %>%
  summarise(observation_count = n())  # Count observations for each drug_count

# Plot the counts with all x-axis values displayed
ggplot(drug_count_summary, aes(x = drug_count, y = observation_count)) +
  geom_bar(stat = "identity", fill = "darkgreen", color = "grey") +  # Bar plot
  geom_text(aes(label = comma(observation_count)), vjust = -0.5, size = 4, color = "black") +  # Add counts above bars
  labs(
    x = "Number of Treatments Used During Labor and Delivery",
    y = "Observation Count"
  ) +
  scale_y_continuous(labels = comma) +  # Format numbers on the y-axis
  scale_x_continuous(breaks = seq(min(drug_count_summary$drug_count), max(drug_count_summary$drug_count), by = 1)) +  # Ensure all x-axis values are shown
  theme_minimal() +
  theme(
    axis.text = element_text(size = 12),
    axis.title = element_text(size = 13)
  )

```

### Type of Treatments Used During Delivery and Labor

```{r}
#| label: fig-typetreat
#| fig-cap: The distribution of observations across different type of treatments used during delivery and labor
#| echo: false
#| warning: false
#| message: false

# Define custom labels and order for Treatment_Type
custom_labels <- c(
  "indc" = "Induction",
  "augmt" = "Augmentation",
  "ster" = "Steroids",
  "antb" = "Antibiotics",
  "chor" = "Chorioamnionitis",
  "anes" = "Anesthesia"
)
# Reshape data to long format for treatment types
treatment_summary <- balanced_data %>%
  select(indc, augmt, ster, antb, chor, anes) %>%  # Select treatment columns
  pivot_longer(
    cols = everything(),  # Reshape all columns to long format
    names_to = "Treatment_Type",  # Column for treatment type
    values_to = "Value"  # Column for values (0 or 1)
  ) %>%
  filter(Value == 1) %>%  # Keep rows where treatment is used
  group_by(Treatment_Type) %>%
  summarise(Count = n(), .groups = "drop") %>%  # Count observations for each treatment type
  mutate(
    Treatment_Type = factor(Treatment_Type, levels = names(custom_labels), labels = custom_labels)  # Apply custom labels and order
  )

# Plot the distribution of observations for each treatment
ggplot(treatment_summary, aes(x = Treatment_Type, y = Count, fill = Treatment_Type)) +
  geom_bar(stat = "identity", fill = "darkgreen", color = "grey") +  # Bar plot
  geom_text(aes(label = comma(Count)), vjust = -0.5, size = 4, color = "black") +  # Add counts above bars
  labs(
    x = "Treatment Type",
    y = "Observation Count"
  ) +
  scale_y_continuous(labels = comma) +  # Format y-axis labels with commas
  theme_minimal() +
  theme(
    axis.text = element_text(size = 10),
    axis.title = element_text(size = 10),
    legend.position = "none"  # Remove legend for cleaner look
  )
```

### Induction of labor

```{r}
#| label: fig-indc1
#| fig-cap: The distribution of observations in the raw dataset based on whether the infant received induction of labor or not
#| echo: false
#| warning: false
#| message: false

# Prepare the data
ld_indl_summary1 <- raw_data %>%
  group_by(ld_indl) %>%
  summarise(Count = n(), .groups = "drop") %>%
  mutate(
    ld_indl = ifelse(ld_indl == "Y", "Yes", ifelse(ld_indl == "N", "No", "Unknown")),  # Label groups
    ld_indl = factor(ld_indl, levels = c("Yes", "No", "Unknown"))  # Specify factor levels for order
  )

# Plot the bar graph
ggplot(ld_indl_summary1, aes(x = ld_indl, y = Count)) +
  geom_bar(stat = "identity", fill = "#a9dfbf", color = "grey") +  # Bar plot
  geom_text(aes(label = comma(Count)), vjust = -0.5, size = 4, color = "black") +  # Add counts above bars
  labs(
    x = "Induction of Labor",
    y = "Observation Count"
  ) +
  scale_y_continuous(limits = c(0, max(ld_indl_summary1$Count) + 1000), expand = c(0, 0)) +
  scale_y_continuous(labels = comma) +  # Format y-axis labels with commas
  theme_minimal() +
  theme(
    axis.text = element_text(size = 11),
    axis.title = element_text(size = 13)
  )

```
```{r}
#| label: fig-indc2
#| fig-cap: The distribution of observations in the analysis dataset on whether the infant received induction of labor or not,
#| echo: false
#| warning: false
#| message: false

# Prepare the data
ld_indl_summary2 <- balanced_data %>%
  group_by(indc) %>%
  summarise(Count = n(), .groups = "drop") %>%
  mutate(
    ld_indl = ifelse(indc == 1, "Yes", ifelse(indc == 0 , "No", "Unknown")),  # Label groups
    ld_indl = factor(ld_indl, levels = c("Yes", "No", "Unknown"))  # Specify factor levels for order
  )

# Plot the bar graph
ggplot(ld_indl_summary2, aes(x = ld_indl, y = Count)) +
  geom_bar(stat = "identity", fill = "#a9dfbf", color = "grey") +  # Bar plot
  geom_text(aes(label = comma(Count)), vjust = -0.5, size = 4, color = "black") +  # Add counts above bars
  labs(
    x = "Induction of Labor",
    y = "Observation Count"
  ) +
  scale_y_continuous(limits = c(0, max(ld_indl_summary2$Count) + 1000), expand = c(0, 0)) +
  scale_y_continuous(labels = comma) +  # Format y-axis labels with commas
  theme_minimal() +
  theme(
    axis.text = element_text(size = 11),
    axis.title = element_text(size = 13)
  )

```

### Augmentation of labor

```{r}
#| label: fig-augmt1
#| fig-cap: The distribution observations in the raw dataset based on whether the infant received augmentation of labor or not
#| echo: false
#| warning: false
#| message: false

# Prepare the data
ld_augm_summary1 <- raw_data %>%
  group_by(ld_augm) %>%
  summarise(Count = n(), .groups = "drop") %>%
  mutate(
    ld_augm = ifelse(ld_augm == "Y", "Yes", ifelse(ld_augm == "N", "No", "Unknown")),  # Label groups
    ld_augm = factor(ld_augm, levels = c("Yes", "No", "Unknown"))  # Specify factor levels for order
  )

# Plot the bar graph
ggplot(ld_augm_summary1, aes(x = ld_augm, y = Count)) +
  geom_bar(stat = "identity", fill = "#a9dfbf", color = "grey") +  # Bar plot
  geom_text(aes(label = comma(Count)), vjust = -0.5, size = 4, color = "black") +  # Add counts above bars
  labs(
    x = "Augmentation of Labor",
    y = "Observation Count"
  ) +
  scale_y_continuous(limits = c(0, max(ld_augm_summary1$Count) + 1000), expand = c(0, 0)) +
  scale_y_continuous(labels = comma) +  # Format y-axis labels with commas
  theme_minimal() +
  theme(
    axis.text = element_text(size = 11),
    axis.title = element_text(size = 13)
  )

```
```{r}
#| label: fig-augmt2
#| fig-cap: The distribution observations in the analysis datset on whether the infant received augmentation of labor or not
#| echo: false
#| warning: false
#| message: false

# Prepare the data
ld_augm_summary2 <- balanced_data %>%
  group_by(augmt) %>%
  summarise(Count = n(), .groups = "drop") %>%
  mutate(
    ld_augm = ifelse(augmt == 1, "Yes", ifelse(augmt == 0 , "No", "Unknown")),  # Label groups
    ld_augm = factor(ld_augm, levels = c("Yes", "No", "Unknown"))  # Specify factor levels for order
  )

# Plot the bar graph
ggplot(ld_augm_summary2, aes(x = ld_augm, y = Count)) +
  geom_bar(stat = "identity", fill = "#a9dfbf", color = "grey") +  # Bar plot
  geom_text(aes(label = comma(Count)), vjust = -0.5, size = 4, color = "black") +  # Add counts above bars
  labs(
    x = "Augmentation of Labor",
    y = "Observation Count"
  ) +
  scale_y_continuous(limits = c(0, max(ld_augm_summary2$Count) + 1000), expand = c(0, 0)) +
  scale_y_continuous(labels = comma) +  # Format y-axis labels with commas
  theme_minimal() +
  theme(
    axis.text = element_text(size = 11),
    axis.title = element_text(size = 13)
  )

```

### Steroids (glucocorticoids) for fetal lung maturation received by the mother before delivery

```{r}
#| label: fig-ster1
#| fig-cap: The distribution observations in the raw dataset based on whether the infant received steroid treatments or not
#| echo: false
#| warning: false
#| message: false

# Prepare the data
ld_ster_summary1 <- raw_data %>%
  group_by(ld_ster) %>%
  summarise(Count = n(), .groups = "drop") %>%
  mutate(
    ld_ster = ifelse(ld_ster == "Y", "Yes", ifelse(ld_ster == "N", "No", "Unknown")),  # Label groups
    ld_ster = factor(ld_ster, levels = c("Yes", "No", "Unknown"))  # Specify factor levels for order
  )

# Plot the bar graph
ggplot(ld_ster_summary1, aes(x = ld_ster, y = Count)) +
  geom_bar(stat = "identity", fill = "#a9dfbf", color = "grey") +  # Bar plot
  geom_text(aes(label = comma(Count)), vjust = -0.5, size = 4, color = "black") +  # Add counts above bars
  labs(
    x = "Steroids Usage",
    y = "Observation Count"
  ) +
  scale_y_continuous(limits = c(0, max(ld_ster_summary1$Count) + 1000), expand = c(0, 0)) +
  scale_y_continuous(labels = comma) +  # Format y-axis labels with commas
  theme_minimal() +
  theme(
    axis.text = element_text(size = 11),
    axis.title = element_text(size = 13)
  )

```
```{r}
#| label: fig-ster2
#| fig-cap: The distribution observations in the analysis datset on whether the infant received steroid treatments or not
#| echo: false
#| warning: false
#| message: false

# Prepare the data
ld_ster_summary2 <- balanced_data %>%
  group_by(ster) %>%
  summarise(Count = n(), .groups = "drop") %>%
  mutate(
    ld_ster = ifelse(ster == 1, "Yes", ifelse(ster == 0 , "No", "Unknown")),  # Label groups
    ld_ster = factor(ld_ster, levels = c("Yes", "No", "Unknown"))  # Specify factor levels for order
  )

# Plot the bar graph
ggplot(ld_ster_summary2, aes(x = ld_ster, y = Count)) +
  geom_bar(stat = "identity", fill = "#a9dfbf", color = "grey") +  # Bar plot
  geom_text(aes(label = comma(Count)), vjust = -0.5, size = 4, color = "black") +  # Add counts above bars
  labs(
    x = "Steroids Usage",
    y = "Observation Count"
  ) +
  scale_y_continuous(limits = c(0, max(ld_ster_summary2$Count) + 1000), expand = c(0, 0)) +
  scale_y_continuous(labels = comma) +  # Format y-axis labels with commas
  theme_minimal() +
  theme(
    axis.text = element_text(size = 11),
    axis.title = element_text(size = 13)
  )

```

### Antibiotics received by the mother during delivery

```{r}
#| label: fig-antb1
#| fig-cap: The distribution observations in the raw dataset based on whether the infant received antibiotic treatments or not
#| echo: false
#| warning: false
#| message: false

# Prepare the data
ld_antb_summary1 <- raw_data %>%
  group_by(ld_antb) %>%
  summarise(Count = n(), .groups = "drop") %>%
  mutate(
    ld_antb = ifelse(ld_antb == "Y", "Yes", ifelse(ld_antb == "N", "No", "Unknown")),  # Label groups
    ld_antb = factor(ld_antb, levels = c("Yes", "No", "Unknown"))  # Specify factor levels for order
  )

# Plot the bar graph
ggplot(ld_antb_summary1, aes(x = ld_antb, y = Count)) +
  geom_bar(stat = "identity", fill = "#a9dfbf", color = "grey") +  # Bar plot
  geom_text(aes(label = comma(Count)), vjust = -0.5, size = 4, color = "black") +  # Add counts above bars
  labs(
    x = "Antibiotics Usage",
    y = "Observation Count"
  ) +
  scale_y_continuous(limits = c(0, max(ld_antb_summary1$Count) + 1000), expand = c(0, 0)) +
  scale_y_continuous(labels = comma) +  # Format y-axis labels with commas
  theme_minimal() +
  theme(
    axis.text = element_text(size = 11),
    axis.title = element_text(size = 13)
  )

```

```{r}
#| label: fig-antb2
#| fig-cap: The distribution observations in the analysis datset on whether the infant received antibiotic treatments or not
#| echo: false
#| warning: false
#| message: false

# Prepare the data
ld_antb_summary2 <- balanced_data %>%
  group_by(antb) %>%
  summarise(Count = n(), .groups = "drop") %>%
  mutate(
    ld_antb = ifelse(antb == 1, "Yes", ifelse(antb == 0 , "No", "Unknown")),  # Label groups
    ld_antb = factor(ld_antb, levels = c("Yes", "No", "Unknown"))  # Specify factor levels for order
  )

# Plot the bar graph
ggplot(ld_antb_summary2, aes(x = ld_antb, y = Count)) +
  geom_bar(stat = "identity", fill = "#a9dfbf", color = "grey") +  # Bar plot
  geom_text(aes(label = comma(Count)), vjust = -0.5, size = 4, color = "black") +  # Add counts above bars
  labs(
    x = "Antibiotics Usage",
    y = "Observation Count"
  ) +
  scale_y_continuous(limits = c(0, max(ld_antb_summary2$Count) + 1000), expand = c(0, 0)) +
  scale_y_continuous(labels = comma) +  # Format y-axis labels with commas
  theme_minimal() +
  theme(
    axis.text = element_text(size = 11),
    axis.title = element_text(size = 13)
  )

```

### Clinical chorioamnionitis diagnosed during labor or maternal temperature over 38 degrees celcius (100.4 degrees fahrenheit)

```{r}
#| label: fig-chor1
#| fig-cap: The distribution observations in the raw dataset based on whether the infant received chorioamnionitis treatments or not
#| echo: false
#| warning: false
#| message: false

# Prepare the data
ld_chor_summary1 <- raw_data %>%
  group_by(ld_chor) %>%
  summarise(Count = n(), .groups = "drop") %>%
  mutate(
    ld_chor = ifelse(ld_chor == "Y", "Yes", ifelse(ld_chor == "N", "No", "Unknown")),  # Label groups
    ld_chor = factor(ld_chor, levels = c("Yes", "No", "Unknown"))  # Specify factor levels for order
  )

# Plot the bar graph
ggplot(ld_chor_summary1, aes(x = ld_chor, y = Count)) +
  geom_bar(stat = "identity", fill = "#a9dfbf", color = "grey") +  # Bar plot
  geom_text(aes(label = comma(Count)), vjust = -0.5, size = 4, color = "black") +  # Add counts above bars
  labs(
    x = "Chorioamnionitis Usage",
    y = "Observation Count"
  ) +
  scale_y_continuous(limits = c(0, max(ld_chor_summary1$Count) + 1000), expand = c(0, 0)) +
  scale_y_continuous(labels = comma) +  # Format y-axis labels with commas
  theme_minimal() +
  theme(
    axis.text = element_text(size = 11),
    axis.title = element_text(size = 13)
  )

```
```{r}
#| label: fig-chor2
#| fig-cap: The distribution observations in the analysis datset on whether the infant received chorioamnionitis treatments or not
#| echo: false
#| warning: false
#| message: false

# Prepare the data
ld_chor_summary2 <- balanced_data %>%
  group_by(chor) %>%
  summarise(Count = n(), .groups = "drop") %>%
  mutate(
    ld_chor = ifelse(chor == 1, "Yes", ifelse(chor == 0 , "No", "Unknown")),  # Label groups
    ld_chor = factor(ld_chor, levels = c("Yes", "No", "Unknown"))  # Specify factor levels for order
  )

# Plot the bar graph
ggplot(ld_chor_summary2, aes(x = ld_chor, y = Count)) +
  geom_bar(stat = "identity", fill = "#a9dfbf", color = "grey") +  # Bar plot
  geom_text(aes(label = comma(Count)), vjust = -0.5, size = 4, color = "black") +  # Add counts above bars
  labs(
    x = "Chorioamnionitis Usage",
    y = "Observation Count"
  ) +
  scale_y_continuous(limits = c(0, max(ld_chor_summary2$Count) + 1000), expand = c(0, 0)) +
  scale_y_continuous(labels = comma) +  # Format y-axis labels with commas
  theme_minimal() +
  theme(
    axis.text = element_text(size = 11),
    axis.title = element_text(size = 13)
  )

```

### Epidural or spinal anesthesia during labor

```{r}
#| label: fig-anes1
#| fig-cap: The distribution observations in the raw dataset based on whether the infant received anesthesia treatments or not
#| echo: false
#| warning: false
#| message: false

# Prepare the data
ld_anes_summary1 <- raw_data %>%
  group_by(ld_anes) %>%
  summarise(Count = n(), .groups = "drop") %>%
  mutate(
    ld_anes = ifelse(ld_anes == "Y", "Yes", ifelse(ld_anes == "N", "No", "Unknown")),  # Label groups
    ld_anes = factor(ld_anes, levels = c("Yes", "No", "Unknown"))  # Specify factor levels for order
  )

# Plot the bar graph
ggplot(ld_anes_summary1, aes(x = ld_anes, y = Count)) +
  geom_bar(stat = "identity", fill = "#a9dfbf", color = "grey") +  # Bar plot
  geom_text(aes(label = comma(Count)), vjust = -0.5, size = 4, color = "black") +  # Add counts above bars
  labs(
    x = "Anesthesia Usage",
    y = "Observation Count"
  ) +
  scale_y_continuous(limits = c(0, max(ld_anes_summary1$Count) + 1000), expand = c(0, 0)) +
  scale_y_continuous(labels = comma) +  # Format y-axis labels with commas
  theme_minimal() +
  theme(
    axis.text = element_text(size = 11),
    axis.title = element_text(size = 13)
  )

```

```{r}
#| label: fig-anes2
#| fig-cap: The distribution observations in the analysis datset on whether the infant received anesthesia treatments or not
#| echo: false
#| warning: false
#| message: false

# Prepare the data
ld_anes_summary2 <- balanced_data %>%
  group_by(anes) %>%
  summarise(Count = n(), .groups = "drop") %>%
  mutate(
    ld_anes = ifelse(anes == 1, "Yes", ifelse(anes == 0 , "No", "Unknown")),  # Label groups
    ld_anes = factor(ld_anes, levels = c("Yes", "No", "Unknown"))  # Specify factor levels for order
  )

# Plot the bar graph
ggplot(ld_anes_summary2, aes(x = ld_anes, y = Count)) +
  geom_bar(stat = "identity", fill = "#a9dfbf", color = "grey") +  # Bar plot
  geom_text(aes(label = comma(Count)), vjust = -0.5, size = 4, color = "black") +  # Add counts above bars
  labs(
    x = "Anesthesia Usage",
    y = "Observation Count"
  ) +
  scale_y_continuous(limits = c(0, max(ld_anes_summary2$Count) + 1000), expand = c(0, 0)) +
  scale_y_continuous(labels = comma) +  # Format y-axis labels with commas
  theme_minimal() +
  theme(
    axis.text = element_text(size = 11),
    axis.title = element_text(size = 13)
  )

```

## Correlation between predictor variables

### Induction of Labor and Augmentation of Labor

```{r}
#| label: fig-corr1
#| fig-cap: Mosaic plot showing the correlation between induction and augmentation of labor
#| echo: false
#| warning: false
#| message: false

# Load necessary library
library(vcd)

# Create a contingency table for "Induction of Labor" and "Augmentation of Labor"
contingency_table <- xtabs(~ indc + augmt, data = balanced_data)

# Perform chi-squared test to get expected frequencies
chi_test <- chisq.test(contingency_table)
expected <- chi_test$expected  # Expected frequencies
residuals <- (contingency_table - expected) / sqrt(expected)  # Compute residuals

# Define a custom shading function manually
custom_shading <- function(residuals) {
  breaks <- c(-Inf, -2, 0, 2, Inf)  # Breaks for shading categories
  colors <- c("#E8F5E9", "#C8E6C9", "#A5D6A7", "#388E3C")  # Light to dark green
  as.character(cut(residuals, breaks = breaks, labels = colors, include.lowest = TRUE))
}

# Assign shading to each cell in the contingency table
color_map <- matrix(custom_shading(as.vector(residuals)), 
                    nrow = nrow(contingency_table), 
                    ncol = ncol(contingency_table))

# Create the mosaic plot with custom colors
mosaic(
  contingency_table,
  gp = gpar(fill = color_map),  # Apply the color map
  labeling_args = list(
    set_varnames = c(indc = "Induction of Labor", augmt = "Augmentation of Labor")  # Custom axis labels
  ))

```

### Usage of Steroids and Antibiotics

```{r}
#| label: fig-corr2
#| fig-cap: Mosaic plot showing the correlation between usage of steroids and antibiotics
#| echo: false
#| warning: false
#| message: false

# Load necessary library
library(vcd)

# Create a contingency table for "Induction of Labor" and "Augmentation of Labor"
contingency_table <- xtabs(~ ster + antb, data = balanced_data)

# Perform chi-squared test to get expected frequencies
chi_test <- chisq.test(contingency_table)
expected <- chi_test$expected  # Expected frequencies
residuals <- (contingency_table - expected) / sqrt(expected)  # Compute residuals

# Define a custom shading function manually
custom_shading <- function(residuals) {
  breaks <- c(-Inf, -2, 0, 2, Inf)  # Breaks for shading categories
  colors <- c("#E8F5E9", "#C8E6C9", "#A5D6A7", "#388E3C")  # Light to dark green
  as.character(cut(residuals, breaks = breaks, labels = colors, include.lowest = TRUE))
}

# Assign shading to each cell in the contingency table
color_map <- matrix(custom_shading(as.vector(residuals)), 
                    nrow = nrow(contingency_table), 
                    ncol = ncol(contingency_table))

# Create the mosaic plot with custom colors
mosaic(
  contingency_table,
  gp = gpar(fill = color_map),  # Apply the color map
  labeling_args = list(
    set_varnames = c(ster = "Steroids", antb = "Antibiotics")  # Custom axis labels
  ))

```

### Usage of Chorioamnionitis and Antibiotics

```{r}
#| label: fig-corr3
#| fig-cap: Mosaic plot showing the correlation between usage of chorioamnionitis and antibiotics
#| echo: false
#| warning: false
#| message: false

# Load necessary library
library(vcd)

# Create a contingency table for "Induction of Labor" and "Augmentation of Labor"
contingency_table <- xtabs(~ chor + antb, data = balanced_data)

# Perform chi-squared test to get expected frequencies
chi_test <- chisq.test(contingency_table)
expected <- chi_test$expected  # Expected frequencies
residuals <- (contingency_table - expected) / sqrt(expected)  # Compute residuals

# Define a custom shading function manually
custom_shading <- function(residuals) {
  breaks <- c(-Inf, -2, 0, 2, Inf)  # Breaks for shading categories
  colors <- c("#E8F5E9", "#C8E6C9", "#A5D6A7", "#388E3C")  # Light to dark green
  as.character(cut(residuals, breaks = breaks, labels = colors, include.lowest = TRUE))
}

# Assign shading to each cell in the contingency table
color_map <- matrix(custom_shading(as.vector(residuals)), 
                    nrow = nrow(contingency_table), 
                    ncol = ncol(contingency_table))

# Create the mosaic plot with custom colors
mosaic(
  contingency_table,
  gp = gpar(fill = color_map),  # Apply the color map
  labeling_args = list(
    set_varnames = c(chor = "Chorioamnionitis", antb = "Antibiotics")  # Custom axis labels
  ))

```

# Model

The goal of our modelling strategy is twofold. Firstly,...

Here we briefly describe the Bayesian analysis model used to investigate... Background details and diagnostics are included in [Appendix -@sec-model-details].

## Model set-up

Define $y_i$ as the number of seconds that the plane remained aloft. Then $\beta_i$ is the wing width and $\gamma_i$ is the wing length, both measured in millimeters.

\begin{align} 
y_i|\mu_i, \sigma &\sim \mbox{Normal}(\mu_i, \sigma) \\
\mu_i &= \alpha + \beta_i + \gamma_i\\
\alpha &\sim \mbox{Normal}(0, 2.5) \\
\beta &\sim \mbox{Normal}(0, 2.5) \\
\gamma &\sim \mbox{Normal}(0, 2.5) \\
\sigma &\sim \mbox{Exponential}(1)
\end{align}

We run the model in R [@citeR] using the `rstanarm` package of @rstanarm. We use the default priors from `rstanarm`.

### Model justification

We expect a positive relationship between the size of the wings and time spent aloft. In particular...

We can use maths by including latex between dollar signs, for instance $\theta$.

# Results

## Results from examining the analysis dataset

```{r}
#| label: fig-result1
#| fig-cap: The mean of APGAR5 scores by the number of treatments recieved during labor and delivery
#| echo: false
#| warning: false
#| message: false

anal_data <- balanced_data %>%
  mutate(
    drug_count = rowSums(select(., c(indc, augmt, ster, antb, chor, anes)) == 1, na.rm = TRUE)
  )

summary_data <- anal_data %>%
  group_by(drug_count) %>%
  summarise(
    mean_apgar5 = mean(apgar5),
    sd_apgar5 = sd(apgar5)
  )

ggplot(summary_data, aes(x = factor(drug_count), y = mean_apgar5)) +
  geom_bar(
    stat = "identity",
    width = 0.5,  # Adjust width for thinner bars
    fill = "skyblue",
    color = "black",
    alpha = 0.7
  ) +
  geom_errorbar(
    aes(ymin = mean_apgar5 - sd_apgar5, ymax = mean_apgar5 + sd_apgar5),
    width = 0.2,  # Width of error bars remains the same
    color = "blue"
  ) +
  geom_text(
    aes(label = round(mean_apgar5, 1)),
    vjust = -0.5,
    size = 4,
    color = "black"
  ) +
  labs(
    x = "Number of Treatments During Labor and Delivery",
    y = "Mean of APGAR5 Score"
  ) +
  scale_y_continuous(
    limits = c(0, max(summary_data$mean_apgar5 + summary_data$sd_apgar5) + 1),
    breaks = seq(0, 10, by = 1)
  ) +
  theme_minimal() +
  theme(
    axis.text = element_text(size = 12),
    axis.title = element_text(size = 13)
  )

```

```{r}
#| label: fig-result2
#| fig-cap: Heatmap of the overall correlation between treatments administered during delivery and APGAR5 scores
#| echo: false
#| warning: false
#| message: false

heatmap_data <- balanced_data %>%
  pivot_longer(
    cols = c(indc, augmt, ster, antb, chor, anes), 
    names_to = "Drug_Usage", 
    values_to = "Usage"
  ) %>%
  filter(Usage == 1) %>%  # Keep rows where drug usage is "Yes" (1)
  group_by(Drug_Usage, apgar5) %>%
  summarise(Count = n(), .groups = "drop") %>%
  pivot_wider(names_from = apgar5, values_from = Count, values_fill = 0)

# Convert the reshaped data back to long format for ggplot
heatmap_long <- heatmap_data %>%
  pivot_longer(-Drug_Usage, names_to = "APGAR5", values_to = "Count") %>%
  mutate(APGAR5 = as.numeric(APGAR5))

# Create heatmap
# Create a mapping of custom labels for drug usage categories
custom_labels <- c(
  indc = "Induction",
  augmt = "Augmentation",
  ster = "Steroids",
  antb = "Antibiotics",
  chor = "Chorioamnionitis",
  anes = "Anesthesia"
)
ggplot(heatmap_long, aes(x = APGAR5, y = Drug_Usage, fill = Count)) +
  geom_tile(color = "white") +
  scale_fill_gradient(
    low = "skyblue", 
    high = "blue", 
    name = "Count of Yes Responses", 
    guide = guide_colorbar(
      name = "Yes Responses",
      barwidth = 1,  # Adjust the width of the legend
      barheight = 6  # Adjust the height of the legend
    )
  ) +
  scale_x_continuous(breaks = 0:10) +  # Show all integers from 0 to 10
  scale_y_discrete(labels = custom_labels) +  # Apply custom labels to y-axis (Drug Usage Categories)
  labs(
    x = "Five-Minute APGAR Score",
    y = "Treatments During Labor and Delivery"
  ) +
  theme_minimal(base_size = 10) +
  theme(
    axis.text.x = element_text(hjust = 1, vjust = 1),  # Angle and align x-axis text
    axis.text.y = element_text(size = 9),  # Increase y-axis text size
    legend.text = element_text(size = 7),  # Reduce legend text size
    legend.title = element_text(size = 7),  # Reduce legend title size
  )

```

```{r}
#| label: fig-indcscore
#| fig-cap: APGAR5 score distribution for the usage of induction labor
#| echo: false
#| warning: false
#| message: false

# Filter data for Induction of Labor (indc == 1)
indc_data <- balanced_data %>% filter(indc == 1)

# Compute the counts for each APGAR5 score
indc_data_summary <- indc_data %>%
  group_by(apgar5) %>%
  summarise(Count = n(), .groups = "drop")

# Plot the bar chart with a trendline
ggplot(indc_data_summary, aes(x = apgar5, y = Count)) +
  geom_col(color = "black", fill = "skyblue", alpha = 0.7) +  # Bar chart
  geom_line(aes(y = Count), color = "red", size = 1.2) +       # Add trendline
  geom_point(color = "red", size = 3) +                        # Highlight points
  labs(
    x = "Five-Minute APGAR Score",
    y = "Observation Count"
  ) +
  theme_minimal() +
  scale_x_continuous(breaks = 0:10) +
  scale_y_continuous(limits = c(0, max(indc_data_summary$Count) + 100), expand = c(0, 0)) +
  theme(
    axis.text.x = element_text(hjust = 1, vjust = 1),
    axis.text.y = element_text(size = 12)
  )

```

```{r}
#| label: fig-augmtscore
#| fig-cap: APGAR5 score distribution for the usage of augmentation labor
#| echo: false
#| warning: false
#| message: false

# Filter data for Augmentation of Labor (augmt == 1)
augmt_data <- balanced_data %>% filter(augmt == 1)

# Compute the counts for each APGAR5 score
augmt_data_summary <- augmt_data %>%
  group_by(apgar5) %>%
  summarise(Count = n(), .groups = "drop")

# Plot the bar chart with a trendline
ggplot(augmt_data_summary, aes(x = apgar5, y = Count)) +
  geom_col(color = "black", fill = "skyblue", alpha = 0.7) +  # Bar chart
  geom_line(aes(y = Count), color = "red", size = 1.2) +       # Add trendline
  geom_point(color = "red", size = 3) +                        # Highlight points
  labs(
    x = "Five-Minute APGAR Score",
    y = "Observation Count"
  ) +
  theme_minimal() +
  scale_x_continuous(breaks = 0:10) +
  scale_y_continuous(limits = c(0, max(augmt_data_summary$Count) + 100), expand = c(0, 0)) +
  theme(
    axis.text.x = element_text(hjust = 1, vjust = 1),
    axis.text.y = element_text(size = 12)
  )

```

```{r}
#| label: fig-sterscore
#| fig-cap: APGAR5 score distribution for the usage of steroids
#| echo: false
#| warning: false
#| message: false

# Filter data for Steroids (ster == 1)
ster_data <- balanced_data %>% filter(ster == 1)

# Compute the counts for each APGAR5 score
ster_data_summary <- ster_data %>%
  group_by(apgar5) %>%
  summarise(Count = n(), .groups = "drop")

# Plot the bar chart with a trendline
ggplot(ster_data_summary, aes(x = apgar5, y = Count)) +
  geom_col(color = "black", fill = "skyblue", alpha = 0.7) +  # Bar chart
  geom_line(aes(y = Count), color = "red", size = 1.2) +       # Add trendline
  geom_point(color = "red", size = 3) +                        # Highlight points
  labs(
    x = "Five-Minute APGAR Score",
    y = "Observation Count"
  ) +
  theme_minimal() +
  scale_x_continuous(breaks = 0:10) +
  scale_y_continuous(limits = c(0, max(ster_data_summary$Count) + 100), expand = c(0, 0)) +
  theme(
    axis.text.x = element_text(hjust = 1, vjust = 1),
    axis.text.y = element_text(size = 12)
  )

```

```{r}
#| label: fig-antbscore
#| fig-cap: APGAR5 score distribution for the usage of antibiotics
#| echo: false
#| warning: false
#| message: false

# Filter data for Antibiotics (antb == 1)
antb_data <- balanced_data %>% filter(antb == 1)

# Compute the counts for each APGAR5 score
antb_data_summary <- antb_data %>%
  group_by(apgar5) %>%
  summarise(Count = n(), .groups = "drop") %>%
  arrange(apgar5)  # Ensure data is sorted by APGAR5

# Plot the bar chart with a trendline
ggplot(antb_data_summary, aes(x = apgar5, y = Count)) +
  geom_col(color = "black", fill = "skyblue", alpha = 0.7) +  # Bar chart
  geom_line(aes(y = Count), color = "red", size = 1.2) +       # Add trendline
  geom_point(color = "red", size = 3) +                        # Highlight points
  labs(
    x = "Five-Minute APGAR Score",
    y = "Observation Count"
  ) +
  theme_minimal() +
  scale_x_continuous(breaks = 0:10) +
  scale_y_continuous(limits = c(0, max(antb_data_summary$Count) + 100), expand = c(0, 0)) +
  theme(
    axis.text.x = element_text(hjust = 1, vjust = 1),
    axis.text.y = element_text(size = 12)
  )

```

```{r}
#| label: fig-chorscore
#| fig-cap: APGAR5 score distribution for the usage of chorioamnionitis
#| echo: false
#| warning: false
#| message: false

# Filter data for Chorioamnionitis (chor == 1)
chor_data <- balanced_data %>% filter(chor == 1)

# Compute the counts for each APGAR5 score
chor_data_summary <- chor_data %>%
  group_by(apgar5) %>%
  summarise(Count = n(), .groups = "drop") %>%
  arrange(apgar5)  # Ensure data is sorted by APGAR5

# Plot the bar chart with a trendline
ggplot(chor_data_summary, aes(x = apgar5, y = Count)) +
  geom_col(color = "black", fill = "skyblue", alpha = 0.7) +  # Bar chart
  geom_line(aes(y = Count), color = "red", size = 1.2) +       # Add trendline
  geom_point(color = "red", size = 3) +                        # Highlight points
  labs(
    x = "Five-Minute APGAR Score",
    y = "Observation Count"
  ) +
  theme_minimal() +
  scale_x_continuous(breaks = 0:10) +
  scale_y_continuous(limits = c(0, max(chor_data_summary$Count) + 50), expand = c(0, 0)) +
  theme(
    axis.text.x = element_text(hjust = 1, vjust = 1),
    axis.text.y = element_text(size = 12)
  )

```

```{r}
#| label: fig-anesscore
#| fig-cap: APGAR5 score distribution for the usage of anesthesia
#| echo: false
#| warning: false
#| message: false

# Filter data for Anesthesia (anes == 1)
anes_data <- balanced_data %>% filter(anes == 1)

# Compute the counts for each APGAR5 score
anes_data_summary <- anes_data %>%
  group_by(apgar5) %>%
  summarise(Count = n(), .groups = "drop") %>%
  arrange(apgar5)  # Ensure data is sorted by APGAR5

# Plot the bar chart with a trendline
ggplot(anes_data_summary, aes(x = apgar5, y = Count)) +
  geom_col(color = "black", fill = "skyblue", alpha = 0.7) +  # Bar chart
  geom_line(aes(y = Count), color = "red", size = 1.2) +       # Add trendline
  geom_point(color = "red", size = 3) +                        # Highlight points
  labs(
    x = "Five-Minute APGAR Score",
    y = "Observation Count"
  ) +
  theme_minimal() +
  scale_x_continuous(breaks = 0:10) +
  scale_y_continuous(limits = c(0, max(anes_data_summary$Count) + 200), expand = c(0, 0)) +
  theme(
    axis.text.x = element_text(hjust = 1, vjust = 1),
    axis.text.y = element_text(size = 12)
  )

```

## Results from the prediction model

Our results are summarized in @tbl-modelresults.

# Discussion

## First discussion point {#sec-first-point}

If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.

## Second discussion point

Please don't use these as sub-heading labels - change them to be what your point actually is.

## Third discussion point

## Weaknesses and next steps

Weaknesses and next steps should also be included.

\newpage

\appendix

# Appendix {-}

# Additional data details

# Model details {#sec-model-details}

## Posterior predictive check

In @fig-ppcheckandposteriorvsprior-1 we implement a posterior predictive check. This shows...

In @fig-ppcheckandposteriorvsprior-2 we compare the posterior with the prior. This shows...

## Diagnostics

@fig-stanareyouokay-1 is a trace plot. It shows... This suggests...

@fig-stanareyouokay-2 is a Rhat plot. It shows... This suggests...

\newpage

# References